\LARGE \bf Mean Estimation from One-Bit Measurements

### 2021-01-06 proof-adaptive-optimality.tex
removed expressions with \pm
ready for submission

We begin with two technical lemmas.
\begin{lem} \label{lem:bound_intervals}
  Let $f$ be a log-concave and symmetric density function
  for which Assumption~\ref{assump:failure_rate} holds. For any $x_1 \geq
  \ldots \geq x_n \in \R$,
  \begin{equation}
    \frac{ \left| \sum_{k=1}^n (-1)^{k+1} f(x_k) \right|^2 }{
      \left( \sum_{k=1}^n (-1)^{k+1} F(x_k) \right) \left(1- \sum_{k=1}^n (-1)^{k+1} F(x_k) \right) } 
    \leq  4f(0)^2.
    \label{eq:lem_bound_intervals}
  \end{equation}
\end{lem}
\begin{lem} \label{lem:fisher_bound}
  Let $X$ be a random variable with a symmetric, log-concave, and
  continuously differentiable density function $f(x)$ such that Assumption~1
  holds. For a Borel measurable set $A$, define
  \begin{equation*}
    B(x) \defeq \begin{cases} 1
      & \mbox{if} ~x \in A, \\
      -1 & \mbox{if}~ x \notin A.
    \end{cases}
  \end{equation*}
  Then the Fisher information of $B$ with respect to $\theta$ is bounded
  from above by $\eta(0)$.
\end{lem}

@@

\begin{proof-of-lemma}[\ref{lem:fisher_bound}]  
  We first note that in the special case where $f$ is a normal density, Lemma~\ref{lem:fisher_bound} follows from \cite[Thm.~3]{Barnes2018}. The proof below, valid for any log-concave symmetric density satisfying Assumption~\ref{assump:failure_rate}, is based on a different techique than that of \cite{Barnes2018}. \\

  Write the Fisher information of $B$ with respect to $\theta$ as
  \begin{align}
    I_\theta & =  \ex{ \left( \frac{d}{d\theta} \log P\left( B | \theta \right) \right)^2 |\theta } \nonumber \\
    & = \frac{ \left(\frac{d}{d\theta} P(B=1|\theta) \right)^2}{P(B=1| \theta)} + \frac{ \left(\frac{d}{d\theta} P(B=-1|\theta) \right)^2} {P(B=-1| \theta)} \nonumber \\
    & =  \frac{ \left( \frac{d}{d\theta} \int_A f \left( x-\theta\right)dx \right)^2} { P(B=1| \theta) } + \frac{ \left( \frac{d}{d\theta}\int_A f \left( x-\theta \right)dx \right)^2} { P(B=-1| \theta) } \nonumber \\ 
    & \overset{(a)}{=} \frac{ \left( - \int_A f' \left( x-\theta \right)dx \right)^2} { P(B=1| \theta) } + \frac{ \left(- \int_A f' \left( x-\theta \right)dx \right)^2} { P(B=-1| \theta) } \nonumber \\ 
    & = \frac{\left( \int_A f'\left( x-\theta \right) dx \right)^2 }{  P(B=1 | \theta) \left(1-P(B=1|\theta) \right)  }, \nonumber \\
    & = \frac{\left( \int_A f'\left( x-\theta \right) dx \right) \left( \int_A f'\left( x-\theta \right) dx \right)}{ \left( \int_A f \left( x-\theta \right) dx \right)  \left(1- \int_A f \left( x-\theta \right) dx \right) }, \label{eq:lem_fisher_bound_proof1}
  \end{align}
  where differentiation under the integral sign in $(a)$ is possible since
  $f$ is log-concave hence a.e.\ differentiable
  (cf.~\cite{Bertsekas73}) with a.e.\ derivative $f'(x)$. Regularity of
  the Lebesgue measure implies that for any $\epsilon>0$, there exists a
  finite number $k$ of disjoint open intervals $I_1,\ldots I_k$ such that
  \begin{equation*}
    \int_{A\setminus \cup_{j=1}^k I_j }  dx < \epsilon.
  \end{equation*}
  It follows that for any $\epsilon' > 0$, the set $A$ in
  \eqref{eq:lem_fisher_bound_proof1} can be replaced by a finite union of
  disjoint intervals without increasing $I_\theta$ by more than
  $\epsilon'$. Consequently, we may proceed assuming that
  $A$ is of the form
  \begin{equation*}
    A = \cup_{j=1}^k (a_j,b_j),
  \end{equation*}
  with $\infty \leq a_1 \leq \ldots a_k$, $b_1 \leq b_k \leq \infty$ and $a_j \leq b_j$ for $j=1,\ldots,k$. Under this assumption we have
  \begin{align*}
    \mathbb P(B_n=1| \theta) & = \sum_{j=1}^k \mathbb P\left(X_n \in (a_j,b_j) \right)  \\
    & = \sum_{j=1}^k \left( F \left(b_j-\theta\right) -  F \left(a_j-\theta\right)  \right),
  \end{align*}
  so we may rewrite Eq.~\eqref{eq:lem_fisher_bound_proof1} as
  \begin{align*}
    I_\theta & =   \frac { \left( \sum_{j=1}^{k} f \left(a_j-\theta \right) - f \left( b_j-\theta \right)  \right)^2 } 
    { \left( \sum_{j=1}^k F \left( b_j-\theta \right) - F \left( a_j-\theta \right)  \right) }  \nonumber \\
    & \times \frac {1} 
    {1- \left( \sum_{j=1}^k F \left(  b_j-\theta \right) - F \left( a_j-\theta \right)  \right) } 
  \end{align*}
  It follows from Lemma~\ref{lem:bound_intervals} that for any $\theta \in
  \R$ and any choice of the intervals' endpoints,
  \begin{equation*}
    I_\theta \le
    \max_{t \in \{a_j,b_j, j=1,\ldots,k\} } 4f(t)^2 \leq 4 f(0)^2, 
  \end{equation*}
  where the last transition is by Assumption~\ref{assump:failure_rate}. 
\end{proof-of-lemma}

We now finish the proof of Theorem~\ref{thm:adpative_lower_bound}.
To bound the Fisher information of any set of $n$ one-bit
messages with respect to $\theta$, we consider the conditional distribution $P({B_1,\ldots,B_n|\theta})$ of
$(B_1,\ldots,B_n)$ given $\theta$. We have
\begin{align*}
  P\left( B_1,\ldots,B_n | \theta \right) & =  \prod_{i=1}^n P\left(B_i | \theta, B_1,\ldots,B_{i-1} \right),
  %% \label{eq:adpt_lower_bound_proof:1}
\end{align*}
where $P\left(B_i =1 | \theta, B_1,\ldots,B_{i-1} \right) = \mathbb P\left(
X_i \in A_i\right)$, when $B_i = \indic{X_i \in A_i}$ by independence,
so that the Fisher information of $B_1,\ldots,B_n$ with
respect to $\theta$ is
\begin{align}
I_\theta(B_1,\ldots,B_N) = \sum_{i=1}^n I_\theta (B_i|B_1,\ldots,B_{i-1}),
\label{eq:fisher_information}
\end{align}
where $I_\theta (B_i|B_{i-1},\ldots,B_1)$ is the Fisher information of the
distribution of $B_i$ given $B_1,\ldots,B_{i-1}$.
Lemma~\ref{lem:fisher_bound} implies that $I_\theta
(B_i|B_{i-1},\ldots,B_1) \leq 4f(0)^2$. The Van Trees inequality
\cite{van2004detection, gill1995applications} now implies
\begin{align*}
\ex{ \left( \theta_n - \theta \right)^2} &  \geq \frac{1}{ \ex{ I_\theta(B_1,\ldots,B_n)} + I_0} \\
& = \frac{1}{ \sum_{i=1}^n I_\theta (B_i | B^{i-1} ) + I_0} \\
& \geq \frac{1}{ 4f^2(0) n + I_0}
\end{align*}
as desired.

@@

\begin{proof-of-lemma}[\ref{lem:bound_intervals_delta}]
  Denote 
  \begin{equation*}
    \delta_n(x_1,\ldots,x_n) \triangleq \sum_{i=1}^{n} s_i f(x_i),
  \end{equation*}
  \begin{equation*}
    \Delta_n(x_1,\ldots,x_n) \triangleq  \sum_{i=1}^n s_i F(x_i), 
  \end{equation*}
  where $s_i \triangleq (-1)^{i+1}$. We use induction on $n \in \N$ to show that 
  \begin{align}
    \label{eq:lemm:interval_bounds:to_show}
    \frac{ \left| \delta_n(x_1,\ldots,x_n) \right|^{2+\delta}} 
         {\left|\Delta_n(x_1,\ldots,x_n)\left(1- \Delta_n(x_1,\ldots,x_n) \right) \right|^{1+\delta} } \leq \max_{i}\eta_{\delta}(x_i).
  \end{align}
  Since 
  \begin{equation*}
    \eta_\delta(x) =  \frac{  \left|\delta_1(x) \right|^{2+\delta}}{\left|\Delta_1(x)
      (1-\Delta_1(x)) \right|^{1+\delta}}, 
  \end{equation*}
  The case $n=1$ is trivial.  
  %follows from the assumption that $\eta_\delta(x)$ is non-increasing in $|x|$. 
  Assume that \eqref{eq:lemm:interval_bounds:to_show} holds for all integers up to $n = N$ and for any $x_1 \ge \ldots \ge x_N$. Consider the case $n = N+1$. Let $i^*$ be the index such that $x_{i^*}$ has minimal absolute value among $x_1,\ldots,x_N$. The assumption on $\eta_\delta(x)$ implies that
  \begin{equation*}
    \eta_\delta(x_{i^*}) = \max_i \eta_\delta(x_i).
  \end{equation*}
  Since the LHS of \eqref{eq:lem_bound_intervals_delta} is invariant to a sign flip of all $x_1,\ldots,x_{N+1}$, we may assume that $x_{i^*}$ is positive without loss of generality. 
  %For simplicity, we also assume that $s_{i^*} = 1$ and note that the case $s_{i^*} = -1$ is obtained using identical arguments. 
  Set $x^* = x_{i^*}$ and let $k=i^*-1$. In what follows, variables with subscript of non-positive index are ignored in summations and in lists of arguments to functions. Consider the function
  \begin{align}
    & g(y_1,\ldots,y_N) \triangleq g(y_1,\ldots,y_N|x^*,k) \\
    &  \triangleq  \frac{\left| \delta_{N+1}(y_1,\ldots,y_k,x^*,y_{k+1}\ldots,y_N) )\right|^{2+\delta}}{
      \left|\Delta_{N+1}(y_1,\ldots,y_k,x_{i^*},y_{k+1}\ldots,y_N)  (1 -\Delta_{N+1}(y_1,\ldots,y_k,x^*,y_{k+1}\ldots,y_N)  \right|^{1+\delta}
    } \label{eq:g_def}.
  \end{align}
  The LHS of \eqref{eq:lemm:interval_bounds:to_show} is obtained by taking $y_i=x_{k_i}$ where $k_i$ is the $i$th element in $\{1,\ldots,N+1\}\setminus \{i^*\}$. It is therefore enough to prove that 
  \begin{equation*}
    \max_{(y_1,\ldots,y_N) \in A_N} g(y_1,\ldots,y_N) \leq \eta_{\delta}(x^*),
  \end{equation*}
  where 
  \begin{equation*}
    A_N(x^*,k) \triangleq \left\{ (y_1,\ldots,y_N) \in \R^N\, : \, y_1 \ge y_k \ge x^* \ge -x^* \ge y_{k+1} \ldots \ge y_N
    \right\}.
  \end{equation*}
  %
  Since $f(x)$ is log-concave and symmetric, we may write $f(x) = e^{c(x)}$
  where $c(x)$ is concave, symmetric, and superdifferentiable on the interior
  of its domain with
  supergradient set $\partial c(x) = \{v \in \R \mid c(y) \le c(x) + v (y - x) ~
  \mbox{for~all}~ y\}$; $c$ is also
  differentiable a.e.\ with derivative
  \begin{equation*}
    c'(x) \triangleq \frac{f'(x)}{f(x)}
  \end{equation*}
  (when it exists), and we otherwise simply treat $f'(x) / f(x) = c'(x) \in
  \partial c(x)$ as an arbitrary element of the superdifferential.  The
  supergradient sets $\partial c(x)$ are increasing, in that $v_0 \in
  \partial c(x_0)$ and $v_1 \in \partial c(x_1)$ implies that $(v_1 - v_0)
  (x_1 - x_0) \le 0$.  We first prove the lemma under the assumption that
  $c$ is strictly concave, or, equivalently, that $v_i \in \partial c(x_i)$
  implies that $(v_1 - v_0)(x_1 - x_0) < 0$ whenever $x_1 \neq x_0$; that
  is, $c'$ is strictly decreasing.

@@

    g(y_1,\ldots,y_N) \leq \eta_{\delta}(x^*)

@@

  \item[(1)]  $y_N \to -\infty$. 
  \item[(2)] $y_1 \to \infty$.
  \item[(3)] $y_k = x_{i^*}$.
  \item[(4)] $y_{k+1} = -x_{i^*}$. 

@@

  For case (1), 

@@

  which is smaller than $\eta_\delta(x_{i^*})$ by the induction hypothesis. Similarly, under case (2),

@@

  which is smaller than $\eta_{\delta}(x_{i^*})$ by the induction hypothesis. Under case (3), the terms in $\delta_N$ and $\Delta_N$ corresponding to $y_k$ and $x_{i^*}$ cancel each other. As a result,  $g(y_1,\ldots,y_N)$ reduces to an expression with $n=N-1$ variables hence this case is handled by the induction hypothesis. 

@@

  Finally, under case (4), set 

>>>>

We begin with two technical lemmas.
\begin{lem} \label{lem:bound_intervals}
  Let $f$ be a log-concave and symmetric density function
  for which Assumption~\ref{assump:failure_rate} holds. For any $x_1 \geq
  \ldots \geq x_n \in \R$,
  \begin{equation}
    \frac{ \left| \sum_{k=1}^n (-1)^{k+1} f(x_k) \right|^2 }{
      \left( \sum_{k=1}^n (-1)^{k+1} F(x_k) \right) \left(1- \sum_{k=1}^n (-1)^{k+1} F(x_k) \right) } 
    \leq  4f(0)^2.
    \label{eq:lem_bound_intervals}
  \end{equation}
\end{lem}
\begin{lem} \label{lem:fisher_bound}
  Let $X$ be a random variable with a symmetric, log-concave, and
  continuously differentiable density function $f(x)$ such that Assumption~1 holds. For a Borel measurable set $A$, define
  \begin{equation*}
    B(x) \defeq \begin{cases} 1
      & \mbox{if} ~x \in A, \\
      -1 & \mbox{if}~ x \notin A.
    \end{cases}
  \end{equation*}
  The Fisher information of $B$ with respect to $\theta$ is bounded from above by $\eta(0)$.
\end{lem}

@@

\begin{proof-of-lemma}[\ref{lem:fisher_bound}]  
  We first note that in the special case where $f$ is a normal density, Lemma~\ref{lem:fisher_bound} follows from \cite[Thm.~3]{Barnes2018}. The proof below, valid for any log-concave symmetric density satisfying Assumption~\ref{assump:failure_rate}, is based on a different techique than that of \cite{Barnes2018}. \par
  Write the Fisher information of $B$ with respect to $\theta$ as
  \begin{align}
    I_\theta & =  \ex{ \left( \frac{d}{d\theta} \log P\left( B | \theta \right) \right)^2 |\theta } \nonumber \\
    & = \frac{ \left(\frac{d}{d\theta} P(B=1|\theta) \right)^2}{P(B=1| \theta)} + \frac{ \left(\frac{d}{d\theta} P(B=-1|\theta) \right)^2} {P(B=-1| \theta)} \nonumber \\
    & =  \frac{ \left( \frac{d}{d\theta} \int_A f \left( x-\theta\right)dx \right)^2} { P(B=1| \theta) } + \frac{ \left( \frac{d}{d\theta}\int_A f \left( x-\theta \right)dx \right)^2} { P(B=-1| \theta) } \nonumber \\ 
    & \overset{(a)}{=} \frac{ \left( - \int_A f' \left( x-\theta \right)dx \right)^2} { P(B=1| \theta) } + \frac{ \left(- \int_A f' \left( x-\theta \right)dx \right)^2} { P(B=-1| \theta) } \nonumber \\ 
    & = \frac{\left( \int_A f'\left( x-\theta \right) dx \right)^2 }{  P(B=1 | \theta) \left(1-P(B=1|\theta) \right)  }, \nonumber \\
    & = \frac{\left( \int_A f'\left( x-\theta \right) dx \right) \left( \int_A f'\left( x-\theta \right) dx \right)}{ \left( \int_A f \left( x-\theta \right) dx \right)  \left(1- \int_A f \left( x-\theta \right) dx \right) }, \label{eq:lem_fisher_bound_proof1}
  \end{align}
  where differentiation under the integral sign in $(a)$ is justified since $f$ is log-concave hence a.e.\ differentiable
  (cf.~\cite{Bertsekas73}) with a.e.\ derivative $f'(x)$. By regularity of the Lebesgue, for any $\epsilon>0$ there exists a finite number $k$ of disjoint open intervals $I_1,\ldots I_k$ such that
  \begin{equation*}
    \int_{A\setminus \cup_{j=1}^k I_j }  dx < \epsilon.
  \end{equation*}
  It follows that for any $\epsilon' > 0$, the set $A$ in
  \eqref{eq:lem_fisher_bound_proof1} can be replaced by a finite union of disjoint intervals without increasing $I_\theta$ by more than
  $\epsilon'$. Consequently, we may proceed assuming that
  $A$ is of the form
  \begin{equation*}
    A = \cup_{j=1}^k (t^+_j,t^-_j),
  \end{equation*}
  with $\infty \leq t^-_1 \leq \ldots t^-_k$, $t^+_1 \leq t^+_k \leq \infty$ and $t^-_j \leq t^+_j$ for $j=1,\ldots,k$. Under this assumption, 
  \begin{align*}
    P_{\theta}(B_n=1) & = \sum_{j=1}^k \left( F \left(t^+_j-\theta\right) -  F \left(t^-_j-\theta\right)  \right),
  \end{align*}
  so we may rewrite Eq.~\eqref{eq:lem_fisher_bound_proof1} as
  \begin{align*}
    I_\theta & = \frac { \left( \sum_{j=1}^{k} \left[ f \left(t^+_j-\theta \right) - f \left( t^-_j-\theta \right) \right] \right)^2 } 
    { \left( \sum_{j=1}^k \left[ F \left( t^+_j-\theta \right) - F \left( t^-_j-\theta \right) \right] \right) }  \nonumber \\
    & \times \frac {1} 
    {1- \left( \sum_{j=1}^k \left[ F \left(  t^+_j-\theta \right) - F \left( t^-_j-\theta \right) \right] \right) } 
  \end{align*}
  It follows from Lemma~\ref{lem:bound_intervals} that for any $\theta \in \R$ and any choice of the intervals' endpoints,
  \begin{equation*}
    I_\theta \le
    \max_{t \in \{t^\pm_1,\ldots,t^\pm_k\} } 4f(t)^2 \leq 4 f(0)^2. 
  \end{equation*}
\end{proof-of-lemma}

We now prove Theorem~\ref{thm:adpative_lower_bound}. Write the Fisher information of $B_1,\ldots,B_n$ with respect to $\theta$ as 
\begin{align*}
I_\theta(B_1,\ldots,B_n) = \sum_{i=1}^n I_\theta (B_i|B_1,\ldots,B_{i-1}),
\end{align*}
where $I_\theta (B_i|B_{i-1},\ldots,B_1)$ is the Fisher information of the distribution of $B_i$ given $B_1,\ldots,B_{i-1}$. By the definition of the adaptive setting, $P_{\theta}(B_i|B_1,\ldots,B_{i-1}) = P_{\theta}(X_i \in A_i)$ for some Borel measurable $A_i$. Consequently, Lemma~\ref{lem:fisher_bound} applies, leading to the bound
\[
 I_\theta (B_i|B_{i-1},\ldots,B_1) \leq 4f(0)^2,
\]
We conclude
\begin{align}
I_\theta(B_1,\ldots,B_n) = 4f(0)^2 n
\label{eq:fisher_information}.
\end{align}
The Van Trees inequality in the version of 
\cite{gill1995applications} holds under the regularity conditions on $\pi(\cdot)$, which implies
\begin{align*}
\ex{ \left( \theta_n - \theta \right)^2} &  \geq \frac{1}{ \E_{\pi} \left[ I_\theta(B_1,\ldots,B_n)\right] + I_0}.
\end{align*}
Combining the last display with \eqref{eq:fisher_information}, we get
\begin{align*}
\ex{ \left( \theta_n - \theta \right)^2} &  \geq \frac{1}{ 4f(0)^2 n + I_0}.
\end{align*}

\cite{van2004detection} and \eqref{eq:fisher_information} imply
\begin{align*}
\ex{ \left( \theta_n - \theta \right)^2} &  \geq \frac{1}{ \ex{ I_\theta(B_1,\ldots,B_n)} + I_0} \\
& = \frac{1}{ \sum_{i=1}^n I_\theta (B_i | B_1,\ldots,B_{i-1}) + I_0} \\
& \geq \frac{1}{ 4f^2(0) n + I_0}.
\end{align*}

@@

\begin{proof-of-lemma}[\ref{lem:bound_intervals_delta}]
  Denote 
  \begin{equation*}
    \delta_n(x_1,\ldots,x_n) \triangleq \sum_{i=1}^{n} s_i f(x_i),
  \end{equation*}
  \begin{equation*}
    \Delta_n(x_1,\ldots,x_n) \triangleq  \sum_{i=1}^n s_i F(x_i), 
  \end{equation*}
  where $s_i \triangleq (-1)^{i+1}$. We use induction on $n \in \N$ to show that 
  \begin{align}
    \label{eq:lemm:interval_bounds:to_show}
    \frac{ \left| \delta_n(x_1,\ldots,x_n) \right|^{2+\delta}} 
         {\left|\Delta_n(x_1,\ldots,x_n)\left(1- \Delta_n(x_1,\ldots,x_n) \right) \right|^{1+\delta} } \leq \max_{i}\eta_{\delta}(x_i).
  \end{align}
  Since 
  \begin{equation*}
    \eta_\delta(x) =  \frac{  \left|\delta_1(x) \right|^{2+\delta}}{\left|\Delta_1(x)
      (1-\Delta_1(x)) \right|^{1+\delta}}, 
  \end{equation*}
  The case $n=1$ is trivial.  
  %follows from the assumption that $\eta_\delta(x)$ is non-increasing in $|x|$. 
  Assume that \eqref{eq:lemm:interval_bounds:to_show} holds for all integers up to $n = N$ and for any $x_1 \ge \ldots \ge x_N$. Consider the case $n = N+1$. Let $i^*$ be the index such that $x_{i^*}$ has minimal absolute value among $x_1,\ldots,x_N$. The assumption on $\eta_\delta(x)$ implies that
  \begin{equation*}
    \eta_\delta(x_{i^*}) = \max_i \eta_\delta(x_i).
  \end{equation*}
  Since the LHS of \eqref{eq:lem_bound_intervals_delta} is invariant to a sign flip of all $x_1,\ldots,x_{N+1}$, we may assume that $x_{i^*}$ is positive without loss of generality. 
  %For simplicity, we also assume that $s_{i^*} = 1$ and note that the case $s_{i^*} = -1$ is obtained using identical arguments. 
  Set $x^* = x_{i^*}$ and let $k=i^*-1$. Consider the function
  \begin{align}
    & g(y_1,\ldots,y_N) \triangleq g(y_1,\ldots,y_N|x^*,k) \\
    &  \triangleq  \frac{\left| \delta_{N+1}(y_1,\ldots,y_k,x^*,y_{k+1}\ldots,y_N) )\right|^{2+\delta}}{
      \left|\Delta_{N+1}(y_1,\ldots,y_k,x_{i^*},y_{k+1}\ldots,y_N)  (1 -\Delta_{N+1}(y_1,\ldots,y_k,x^*,y_{k+1}\ldots,y_N)  \right|^{1+\delta}
    } \label{eq:g_def}.
  \end{align}
  The LHS of \eqref{eq:lemm:interval_bounds:to_show} is obtained by taking $y_i=x_{k_i}$ where $k_i$ is the $i$th element in $\{1,\ldots,N+1\}\setminus \{i^*\}$. It is therefore enough to prove that 
  \begin{equation*}
    \max_{(y_1,\ldots,y_N) \in A_N(x^*,k) } g(y_1,\ldots,y_N) \leq \eta_{\delta}(x^*),
  \end{equation*}
  where 
  \begin{equation*}
    A_N(x^*,k) \triangleq \left\{ (y_1,\ldots,y_N) \in \R^N\, : \, y_1 \ge y_k \ge x^* \ge -x^* \ge y_{k+1} \ldots \ge y_N
    \right\}.
  \end{equation*}
  %
  Since $f(x)$ is log-concave and symmetric, we may write $f(x) = e^{c(x)}$
  where $c(x)$ is concave, symmetric, and superdifferentiable on the interior
  of its domain with
  supergradient set $\partial c(x) = \{v \in \R \mid c(y) \le c(x) + v (y - x) ~
  \mbox{for~all}~ y\}$; $c$ is also
  differentiable a.e.\ with derivative
  \begin{equation*}
    c'(x) \triangleq \frac{f'(x)}{f(x)}
  \end{equation*}
  (when it exists), and we otherwise simply treat $f'(x) / f(x) = c'(x) \in
  \partial c(x)$ as an arbitrary element of the superdifferential.  The
  supergradient sets $\partial c(x)$ are increasing, in that $v_0 \in
  \partial c(x_0)$ and $v_1 \in \partial c(x_1)$ implies that $(v_1 - v_0)
  (x_1 - x_0) \le 0$.  We first prove the lemma under the assumption that
  $c$ is strictly concave, or, equivalently, that $v_i \in \partial c(x_i)$
  implies that $(v_1 - v_0)(x_1 - x_0) < 0$ whenever $x_1 \neq x_0$; that
  is, $c'$ is strictly decreasing.

@@

    g(y_1,\ldots,y_N) \leq \eta_{\delta}(x^*).

@@

  \item[(i)]  $y_N = -\infty$. 
  \item[(ii)] $y_1 = \infty$.
  \item[(iii)] $y_k = x_{i^*}$.
  \item[(iv)] $y_{k+1} = -x_{i^*}$. 

@@

  For case (i), 

@@

  which is smaller than $\eta_\delta(x_{i^*})$ by the induction hypothesis. Similarly, under case (ii),

@@

  which is smaller than $\eta_{\delta}(x_{i^*})$ by the induction hypothesis. Under case (iii), the terms in $\delta_N$ and $\Delta_N$ corresponding to $y_k$ and $x_{i^*}$ cancel each other. As a result,  $g(y_1,\ldots,y_N)$ reduces to an expression with $N-2$ variables hence this case is handled by the induction hypothesis. 

@@

  Finally, under case (iv), set 
